---
title: "R Analysis"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    mathjax: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,
  message = FALSE,
  warning = FALSE
)
```

## Overview & Setup

This document reproduces the mixed-effects analyses reported in the paper.
Participant-level data files are **not** included in this repository for privacy reasons.

```{r libraries}
library(lme4)
library(lmerTest)
library(emmeans)
library(readr)

# Update this path to wherever you store the analysis CSVs locally.
data_dir <- "PATH/TO/YOUR/DATA"

ns_data_path  <- file.path(data_dir, "ns_data.csv")
nns_data_path <- file.path(data_dir, "nns_data.csv")
mcq_data_path <- file.path(data_dir, "mcq_data.csv")

emm_options(pbkrtest.limit = 3744)  # matches total number of observations
```

---

## NS-Rated Communicative Competence

Four dimensions rated by NS evaluators on a 10-point Likert scale (higher = better):

| Variable | Dimension |
|----------|-----------|
| `ns_wf_comb` | Well-formedness |
| `ns_ca`      | Contextual appropriateness |
| `ns_u1`      | Understandability (NS) |
| `ns_u2`      | Understandability (NNS) |

```{r rq1-ns-load}
ns_df <- read_csv(ns_data_path)
ns_df$condition <- relevel(as.factor(ns_df$condition), ref = "control")
ns_df$language  <- relevel(as.factor(ns_df$language),  ref = "Chinese")

ns_df$ns_wf_comb <- (ns_df$ns_accept + ns_df$ns_wf) / 2
```

### Mixed-effects models

Fixed effects: `condition * language`, `nns_familiarity`, `social_media_usage`.
Random effects: by-participant (`nns_part_id`) and by-NS-evaluator (`ns_eval_id`).

```{r rq1-ns-models}
model_wf_comb <- lmer(
  ns_wf_comb ~ condition * language + nns_familiarity + social_media_usage +
    (1 | nns_part_id) + (1 | ns_eval_id),
  data = ns_df
)

model_ca <- lmer(
  ns_ca ~ condition * language + nns_familiarity + social_media_usage +
    (1 | nns_part_id) + (1 | ns_eval_id),
  data = ns_df
)

model_u1 <- lmer(
  ns_u1 ~ condition * language + nns_familiarity + social_media_usage +
    (1 | nns_part_id) + (1 | ns_eval_id),
  data = ns_df
)

model_u2 <- lmer(
  ns_u2 ~ condition * language + nns_familiarity + social_media_usage +
    (1 | nns_part_id) + (1 | ns_eval_id),
  data = ns_df
)

summary(model_wf_comb)
summary(model_ca)
summary(model_u1)
summary(model_u2)
```

### Post-hoc pairwise comparisons

Bonferroni-adjusted pairwise contrasts applied to all four dimensions.

```{r rq1-ns-emmeans}
for (model in list(model_wf_comb, model_ca, model_u1, model_u2)) {
  emm <- emmeans(model, ~ condition)
  print(emm)
  print(pairs(emm, adjust = "bonferroni"))
}
```

---

## NNS Self-Reported Perceptions

Confidence and helpfulness rated on a 5-point Likert scale (higher = better).
Helpfulness was not collected for the Control condition.

| Variable | Dimension |
|----------|-----------|
| `confidence`        | Confidence (Learning) |
| `confidence_page2`  | Confidence (Production) |
| `helpfulness`       | Helpfulness (Learning) |
| `helpfulness_page2` | Helpfulness (Production) |

```{r rq2-load}
conf_df <- read_csv(nns_data_path)
conf_df$condition <- relevel(as.factor(conf_df$condition), ref = "control")
conf_df$language  <- relevel(as.factor(conf_df$language),  ref = "Chinese")

conf_df_no_ctrl <- conf_df[conf_df$condition != "control", ]
```

### Mixed-effects models

Fixed effects: `condition * language`, `familiarity_rating`, `social_media_usage`.
Random effects: by-participant (`pid`) and by-word (`word`).

```{r rq2-models}
model_conf_learn <- lmer(
  confidence ~ condition * language + familiarity_rating + social_media_usage +
    (1 | pid) + (1 | word),
  data = conf_df
)

model_conf_prod <- lmer(
  confidence_page2 ~ condition * language + familiarity_rating + social_media_usage +
    (1 | pid) + (1 | word),
  data = conf_df
)

model_help_learn <- lmer(
  helpfulness ~ condition * language + familiarity_rating + social_media_usage +
    (1 | pid) + (1 | word),
  data = conf_df_no_ctrl
)

model_help_prod <- lmer(
  helpfulness_page2 ~ condition * language + familiarity_rating + social_media_usage +
    (1 | pid) + (1 | word),
  data = conf_df_no_ctrl
)

summary(model_conf_learn)
summary(model_conf_prod)
summary(model_help_learn)
summary(model_help_prod)
```

### Post-hoc pairwise comparisons

```{r rq2-emmeans}
for (model in list(model_conf_learn, model_conf_prod, model_help_learn, model_help_prod)) {
  emm <- emmeans(model, ~ condition)
  print(emm)
  print(pairs(emm, adjust = "bonferroni"))
}
```

---

## NNS Comprehension Distance

For each writing sample shown to NNS participants, distance is computed as
$|{c} - (c_1^* + c_2^*) / 2|$, where ${c}$ is the NNS rating and $c_1^*, c_2^*$ are NS evaluator ratings collected prior to the study for each writing sample. Lower = better.

```{r rq3-mcq-load}
mcq_df <- read_csv(mcq_data_path)
mcq_df$condition <- relevel(as.factor(mcq_df$condition), ref = "control")
mcq_df$language  <- relevel(as.factor(mcq_df$language),  ref = "Chinese")
```

### Mixed-effects models

```{r rq3-mcq-models}
model_mcq_center <- lmer(
  abs_center_distance ~ condition * language + familiarity_rating + social_media_usage +
    (1 | pid) + (1 | word),
  data = mcq_df
)
summary(model_mcq_center)
```

### Post-hoc pairwise comparisons

```{r rq3-mcq-emmeans}
emm_mcq_center <- emmeans(model_mcq_center, ~ condition)
print(emm_mcq_center)
pairs(emm_mcq_center, adjust = "bonferroni")
```
